The role of cognateness in non-native spoken word recognition

### 2022-02-18 manuscript/manuscript.Rmd
Render README as target

abstract: |

@@

coefs <- fixef(model_fits$fit_6) %>% 
  as.data.frame() %>% 
  rownames_to_column("variable") %>% 
  clean_names() %>% 
  mutate_at(vars(estimate:q97_5), ~ifelse(variable=="Intercept", inv_logit_scaled(.), ./4)) %>% 
  group_split(variable) %>% 
  set_names(make_clean_names(map(., "variable"))) %>% 
  map(select, -variable) %>% 
  map(unlist)

str_repl <- c(
  "Intercept" = "Intercept",
  "frequency_zipf" = "Frequency\n[+1 SD]",
  "pthn:lv" = "PTHN \u00d7 Levenshtein", 
  "lv:group1" = "Levenshtein \u00d7 Group 1",
  "lv:group2" = "Levenshtein \u00d7 Group 2",
  "lv" = "Levenshtein\n[+1 SD]",
  "pthn" = "PTHN\n[+1 SD]",
  "group1" =  "Group 1\n[ENG-* vs. SPA-CAT]",
  "group2" = "Group 2\n[ENG-CAT vs. ENG-SPA]"
)

@@

# Pearson correlations between random effects
corr_mat <- VarCorr(model_fits$fit_6, summary = TRUE)$participant$cor 
corr_names <- rownames(VarCorr(model_fits$fit_6, summary = TRUE)$participant$cor)

@@

```{r bayesfactors}
h_frequency <- abs(hypothesis(model_fits$fit_6, hypothesis = "frequency_zipf = 0")$hypothesis$Evid.Ratio)
h_pthn <- abs(hypothesis(model_fits$fit_6, hypothesis = "pthn = 0")$hypothesis$Evid.Ratio)
h_lv <- abs(hypothesis(model_fits$fit_6, hypothesis = "lv = 0")$hypothesis$Evid.Ratio)
h_pthn_lv <- abs(hypothesis(model_fits$fit_6, hypothesis = "pthn:lv = 0")$hypothesis$Evid.Ratio)
h_group1 <- abs(hypothesis(model_fits$fit_6, hypothesis = "group1 = 0")$hypothesis$Evid.Ratio)
h_group2 <- abs(hypothesis(model_fits$fit_6, hypothesis = "group1 = 0")$hypothesis$Evid.Ratio)
h_lv_group1 <- abs(hypothesis(model_fits$fit_6, hypothesis = "lv:group1 = 0")$hypothesis$Evid.Ratio)
h_lv_group2 <- abs(hypothesis(model_fits$fit_6, hypothesis = "lv:group2 = 0")$hypothesis$Evid.Ratio)

icc_conditional <- icc(model_fits$fit_6, by_group = TRUE)$ICC[2]

sds <- gather_draws(model_fits$fit_6, `sd_.*`, regex = TRUE) %>% 
  mean_qi() %>% 
  mutate(
    .value = ifelse(grepl("Intercept", .variable), inv_logit_scaled(.value), .value/4),
    .lower = ifelse(grepl("Intercept", .variable), inv_logit_scaled(.lower), .lower/4),
    .upper = ifelse(grepl("Intercept", .variable), inv_logit_scaled(.upper), .upper/4)
  ) %>% 
  split(.$.variable)

@@

word_example <- ranef(model_fits$fit_6)$word[,,"Intercept"] %>% 
  as.data.frame() %>% 
  rownames_to_column("word") %>% 
  arrange(-Estimate) %>% 
  clean_names() %>% 
  mutate_if(is.numeric, inv_logit_scaled) %>% 
  slice(1, n()) %>% 
  left_join(distinct(stimuli, word1, ipa1), by = c("word" = "word1")) %>% 
  split(.$word)

@@

cors_range <- gather_draws(model_fits$fit_6, `cor_.*`, regex = TRUE) %>% 
  median_qi() %>% 
  summarise(.value = range(.value))

corr_mat <- VarCorr(model_fits$fit_6, summary = TRUE)$participant$cor 
corr_names <- rownames(VarCorr(model_fits$fit_6, summary = TRUE)$participant$cor)

@@

h_cor_intercept_frequency <- hypothesis(model_fits$fit_6, "participant__Intercept__frequency_zipf = 0", class = "cor")$hypothesis$Evid.Ratio
h_cor_frequency_zipf_lv_group1 <- hypothesis(model_fits$fit_6, "participant__frequency_zipf__lv:group1 = 0", class = "cor")$hypothesis$Evid.Ratio
h_cor_intercept_pthn <- hypothesis(model_fits$fit_6, "participant__Intercept__pthn = 0", class = "cor")$hypothesis$Evid.Ratio
```

Overall, participants were `r percent(coefs$intercept["estimate"], accuracy = 0.01)`, (95% *CrI* = [`r percent(coefs$intercept["q2_5"], accuracy = 0.01)`, `r percent(coefs$intercept["q97_5"], accuracy = 0.01)`]) likely to produce correct translations. The lexical frequency of the correct translation had little impact in accuracy: every standard deviation increment (*SD* = `r printnum(sd(stimuli$frequency_zipf, na.rm = TRUE))` Zipf points) decreased the probability of correct translation in `r percent(coefs$frequency_zipf["estimate"], accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$frequency_zipf["estimate"], accuracy = 0.01)`, `r percent(coefs$frequency_zipf["q2_5"], accuracy = 0.01)`], $BF_{\beta\neq< 0}$ = `r printnum(h_frequency)`). The number of phonological neighbours of the correct translation with higher lexical frequency (*PTHN*) slightly decreased the probability of a correct responses in `r percent(coefs$pthn["estimate"], accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$pthn["q2_5"], accuracy = 0.01)`, `r percent(coefs$pthn["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_pthn)`) for every increase standard deviation increment  (1 *SD* = `r printnum(sd(stimuli$pthn, na.rm = TRUE))` neighbours). Finally, phonological similarity between translations had a strong effect on participants' accuracy: for every *SD* increment in Levenshtein similarity (1 *SD* = `r printnum(percent(sd(stimuli$lv, na.rm = TRUE)), digits = 2)` similarity), translation accuracy increased in `r percent(coefs$lv["estimate"], accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$lv["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = $>3.09 \times 10^{18}$). This difference increased in `r percent(coefs$pthn_lv["estimate"], accuracy = 0.01)` for every *SD* increment in *PTHN* (95% *CrI* = [`r percent(coefs$pthn_lv["q2_5"], accuracy = 0.01)`, `r percent(coefs$pthn_lv["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_pthn_lv)`)

Spanish natives translating Catalan words were `r percent(coefs$group1["estimate"], accuracy = 0.01)` more accurate than English natives translating Spanish and Catalan words (95% *CrI* = [`r percent(coefs$group1["q2_5"], accuracy = 0.01)`, `r percent(coefs$group1["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_group1)`). This difference increased in `r percent(coefs$lv_group1["estimate"], accuracy = 0.01)` for every one *SD* increment in Levenshtein similarity (95% *CrI* = [`r percent(coefs$lv_group1["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv_group1["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_lv_group1)`). English natives translating Catalan words were `r percent(abs(coefs$group2["estimate"]), accuracy = 0.01)` more accurate than those translating Spanish words (95% *CrI* = [`r percent(coefs$group2["q2_5"], accuracy = 0.01)`, `r percent(coefs$group2["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_group2)`). For every one *SD* increment in Levenshtein similarity, the effect increased in `r percent(abs(coefs$lv_group2["estimate"]), accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$lv_group2["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv_group2["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_lv_group1)`).

@@

m <- posterior_epreds_fixed %>% 
  mutate_at(vars(pthn, frequency_zipf), function(x) paste0(x, " SD")) %>% 
  mutate(
    pthn = case_when(
      pthn=="-1 SD" ~ "Low PTHN",
      pthn=="0 SD" ~ "Mean PTHN",
      pthn=="1 SD" ~ "High PTHN"
    ),
    frequency_zipf = case_when(
      frequency_zipf=="-1 SD" ~ "Low frequency",
      frequency_zipf=="0 SD" ~ "Mean frequency",
      frequency_zipf=="1 SD" ~ "High frequency"
    )
  )

ggplot(m, aes(lv, .epred, colour = pthn, fill = pthn)) +
  facet_wrap(~group) +
  geom_hline(yintercept = 0.5, colour = "grey") +
  geom_line(aes(group = interaction(pthn, .draw)), alpha = 0.25, size = 0.5, show.legend = TRUE) +
  stat_summary(aes(linetype = pthn), fun = mean, geom = "line", colour = "black", size = 0.75, show.legend = TRUE) +
  annotate("segment", x = -0.5, xend = -0.9, y = 0.85, yend = 0.85, arrow = arrow(length = unit(0.1, "cm")), size = .5) +
  annotate("segment", x = 2.25, xend = 2.65, y = 0.85, yend = 0.85, arrow = arrow(length = unit(0.1, "cm")), size = .5) +
  annotate("text", x = -0.5, y = 0.95, size = 2.75, label = "Less similar") +
  annotate("text", x = 2.15, y = 0.95, size = 2.75, label = "More similar") +
  labs(x = "Phonological similarity\n(Levenshtein, SD)", y = "P(Correct)", 
       colour = "PTHN", fill = "PTHN", linetype = "PTHN",
       title = "Population-level predictions") +
  scale_color_manual(values = c("#1E88E5", "#ff2976")) +
  scale_y_continuous(labels = function(x) percent(round(x, 2)), limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
  theme(
    legend.position = "top",
    plot.title = element_blank()
  )

@@

The conditional intra-class correlation (*ICC*) for items was `r percent(icc_conditional, accuracy = 0.01)`, indicating that responses from different participants to the same item were, on average, more similar to each other than responses to other items [@gelman2006data]. Item-level intercepts showed remarkable variability (*SD* = `r percent(sds$sd_word_intercept$.value, accuracy = 0.01)`, 95% CrI = [`r percent(sds$sd_word_intercept$.lower, accuracy = 0.01)`, `r percent(sds$sd_word_intercept$.upper, accuracy = 0.01)`]), indicating that participants' accuracy differed substantially across words: some words were correctly translated nearly always (e.g., Spanish *tomate*-*tomato*, /tomate/-/təmɑːtoʊ/, `r percent(word_example$tomate$estimate, accuracy = 0.01)`, 95% CrI = [`r percent(word_example$tomate$q2_5, accuracy = 0.01)`, `r percent(word_example$tomate$q97_5, accuracy = 0.01)`]), while others were almost never translated correctly (e.g., Catalan *colom*-*pigeon*, /kulom/-/pɪdʒɪn/, `r percent(word_example$colom$estimate, accuracy = 0.01)`, 95% CrI = [`r percent(word_example$colom$q2_5, accuracy = 0.01)`, `r percent(word_example$colom$q97_5, accuracy = 0.01)`]). Participant-level intercepts also showed high variability (*SD* = `r percent(sds$sd_participant_intercept$.value, accuracy = 0.01)`, 95% CrI = [`r percent(sds$sd_participant_intercept$.lower, accuracy = 0.01)`, `r percent(sds$sd_participant_intercept$.upper, accuracy = 0.01)`]), suggesting that average translation accuracy also varied significantly across participants. Finally, participant-level estimated slopes showed less variability: *SD* of $Frequency$, $PTHN$, $Levenshtein$, $PTHN \times Levenshtein$, and $Levenshtein \times Group 2$ ranged between `r percent(participant_ranef_range[1], accuracy = 0.01)`, and `r percent(participant_ranef_range[2], accuracy = 0.01)`. The participant-level estimated slopes of the $Leveshtein \times Group 1$ interaction are an exception, though (*SD* = `r percent(sds$sd_participant_lv_group1$.value, accuracy = 0.01)`, 95% CrI = [`r percent(sds$sd_participant_lv_group1$.lower, accuracy = 0.01)`, `r percent(sds$sd_participant_lv_group1$.upper, accuracy = 0.01)`]). This indicates that the impact of phonological similarity ($Levenshtein$) on the difference in accuracy between English and Spanish natives differed substantially across participants.

Finally, participant-level effects showed weak Pearson correlations overall (<10%), with three exceptions. The estimated correlation between the intercepts and the slopes of $Frequency$ was `r percent(corr_data$intercept$estimate_frequency_zipf, accuracy = 0.01)` (95% CrI = [`r percent(corr_data$intercept$q2_5_frequency_zipf, accuracy = 0.01)`, `r percent(corr_data$intercept$q97_5_frequency_zipf, accuracy = 0.01)`], $BF_{Cor \neq 0}$ = `r `round(h_cor_intercept_frequency, 2)`). The estimated correlation between the slopes of $Frequency$ and $Levenshtein \times Group 1$ was `r percent(corr_data$frequency_zipf$estimate_lv_group1, accuracy = 0.01)` (95% CrI = [`r percent(corr_data$frequency_zipf$q2_5_lv_group1, accuracy = 0.01)`, `r percent(corr_data$frequency_zipf$q97_5_lv_group1, accuracy = 0.01)`], $BF_{Cor \neq 0}$ = `r round(h_cor_frequency_zipf_lv_group1, 2)`). The estimated correlation between the intercepts and the slopes of $PTHN$ was `r percent(corr_data$intercept$estimate_pthn, accuracy = 0.01)` (95% CrI = [`r percent(corr_data$intercept$q2_5_pthn, accuracy = 0.01)`, `r percent(corr_data$intercept$q97_5_pthn, accuracy = 0.01)`], $BF_{Cor \neq 0}$ = `r round(h_cor_intercept_pthn, 2)`). The large uncertainty associated with these estimates, however, prevents us from interpreting the size of these correlations as practically meaningful.

>>>>

abstract: |
  Understanding spoken words in a non-native language is easier when their translations to the native language are phonologically similar.

@@

coefs <- fixef(fit_6) %>% 
  as.data.frame() %>% 
  rownames_to_column("variable") %>% 
  clean_names() %>% 
  mutate_at(vars(estimate:q97_5), ~ifelse(variable=="Intercept", inv_logit_scaled(.), ./4)) %>% 
  group_split(variable) %>% 
  set_names(make_clean_names(map(., "variable"))) %>% 
  map(select, -variable) %>% 
  map(unlist)

str_repl <- c(
  "Intercept" = "Intercept",
  "frequency_zipf_std" = "Frequency\n[+1 SD]",
  "pthn_std:lv_std" = "PTHN \u00d7 Levenshtein", 
  "lv_std:groupENGvs.SPA" = "Levenshtein \u00d7 Group 1",
  "lv_std:groupENGMCATvs.ENGMSPA" = "Levenshtein \u00d7 Group 2",
  "lv_std" = "Levenshtein\n[+1 SD]",
  "pthn_std" = "PTHN\n[+1 SD]",
  "groupENGvs.SPA" =  "Group 1\n[ENG-* vs. SPA-CAT]",
  "groupENGMCATvs.ENGMSPA" = "Group 2\n[ENG-CAT vs. ENG-SPA]"
)

@@

# Pearson correlations between random effects
corr_mat <- VarCorr(fit_6, summary = TRUE)$participant$cor 
corr_names <- rownames(VarCorr(fit_6, summary = TRUE)$participant$cor)

@@

```{r bayesfactors}
h_frequency <- abs(hypothesis(fit_6, hypothesis = "frequency_zipf_std = 0")$hypothesis$Evid.Ratio)
h_pthn <- abs(hypothesis(fit_6, hypothesis = "pthn_std = 0")$hypothesis$Evid.Ratio)
h_lv <- abs(hypothesis(fit_6, hypothesis = "lv_std = 0")$hypothesis$Evid.Ratio)
h_pthn_lv <- abs(hypothesis(fit_6, hypothesis = "pthn_std:lv_std = 0")$hypothesis$Evid.Ratio)
h_group1 <- abs(hypothesis(fit_6, hypothesis = "groupENGvs.SPA = 0")$hypothesis$Evid.Ratio)
h_group2 <- abs(hypothesis(fit_6, hypothesis = "groupENGMCATvs.ENGMSPA = 0")$hypothesis$Evid.Ratio)
h_lv_group1 <- abs(hypothesis(fit_6, hypothesis = "lv_std:groupENGvs.SPA = 0")$hypothesis$Evid.Ratio)
h_lv_group2 <- abs(hypothesis(fit_6, hypothesis = "lv_std:groupENGMCATvs.ENGMSPA = 0")$hypothesis$Evid.Ratio)

icc_conditional <- icc(fit_6, by_group = TRUE)$ICC[2]

sds <- gather_draws(fit_6, `sd_.*`, regex = TRUE) %>% 
  mean_qi() %>% 
  mutate(
    .value = ifelse(grepl("Intercept", .variable), inv_logit_scaled(.value), .value/4),
    .lower = ifelse(grepl("Intercept", .variable), inv_logit_scaled(.lower), .lower/4),
    .upper = ifelse(grepl("Intercept", .variable), inv_logit_scaled(.upper), .upper/4)
  ) %>% 
  split(.$.variable)

@@

word_example <- ranef(fit_6)$word[,,"Intercept"] %>% 
  as.data.frame() %>% 
  rownames_to_column("word") %>% 
  arrange(-Estimate) %>% 
  clean_names() %>% 
  mutate_if(is.numeric, inv_logit_scaled) %>% 
  slice(1, n()) %>% 
  left_join(distinct(stimuli, word1, ipa1), by = c("word" = "word1")) %>% 
  split(.$word)

@@

cors_range <- gather_draws(fit_6, `cor_.*`, regex = TRUE) %>% 
  median_qi() %>% 
  summarise(.value = range(.value))

corr_mat <- VarCorr(fit_6, summary = TRUE)$participant$cor 
corr_names <- rownames(VarCorr(fit_6, summary = TRUE)$participant$cor)

@@

h_cor_intercept_frequency <- hypothesis(fit_6, "participant__Intercept__frequency_zipf_std = 0", class = "cor")$hypothesis$Evid.Ratio
h_cor_frequency_zipf_lv_group1 <- hypothesis(fit_6, "participant__frequency_zipf_std__lv_std:groupENGvs.SPA = 0", class = "cor")$hypothesis$Evid.Ratio
h_cor_intercept_pthn <- hypothesis(fit_6, "participant__Intercept__pthn_std = 0", class = "cor")$hypothesis$Evid.Ratio
```

Overall, participants were `r percent(coefs$intercept["estimate"], accuracy = 0.01)`, (95% *CrI* = [`r percent(coefs$intercept["q2_5"], accuracy = 0.01)`, `r percent(coefs$intercept["q97_5"], accuracy = 0.01)`]) likely to produce correct translations. The lexical frequency of the correct translation had little impact in accuracy: every standard deviation increment (*SD* = `r printnum(sd(stimuli$frequency_zipf, na.rm = TRUE))` Zipf points) decreased the probability of correct translation in `r percent(coefs$frequency_zipf_std["estimate"], accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$frequency_zipf_std["estimate"], accuracy = 0.01)`, `r percent(coefs$frequency_zipf_std["q2_5"], accuracy = 0.01)`], $BF_{\beta\neq< 0}$ = `r printnum(h_frequency)`). The number of phonological neighbours of the correct translation with higher lexical frequency (*PTHN*) slightly decreased the probability of a correct responses in `r percent(coefs$pthn_std["estimate"], accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$pthn_std["q2_5"], accuracy = 0.01)`, `r percent(coefs$pthn_std["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_pthn)`) for every increase standard deviation increment  (1 *SD* = `r printnum(sd(stimuli$pthn, na.rm = TRUE))` neighbours). Finally, phonological similarity between translations had a strong effect on participants' accuracy: for every *SD* increment in Levenshtein similarity (1 *SD* = `r printnum(percent(sd(stimuli$lv, na.rm = TRUE)), digits = 2)` similarity), translation accuracy increased in `r percent(coefs$lv_std["estimate"], accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$lv_std["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv_std["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = $>3.09 \times 10^{18}$). This difference increased in `r percent(coefs$pthn_std_lv_std["estimate"], accuracy = 0.01)` for every *SD* increment in *PTHN* (95% *CrI* = [`r percent(coefs$pthn_std_lv_std["q2_5"], accuracy = 0.01)`, `r percent(coefs$pthn_std_lv_std["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_pthn_lv)`)

Spanish natives translating Catalan words were `r percent(coefs$group_en_gvs_spa["estimate"], accuracy = 0.01)` more accurate than English natives translating Spanish and Catalan words (95% *CrI* = [`r percent(coefs$group_en_gvs_spa["q2_5"], accuracy = 0.01)`, `r percent(coefs$group_en_gvs_spa["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_group1)`). This difference increased in `r percent(coefs$lv_std_group_en_gvs_spa["estimate"], accuracy = 0.01)` for every one *SD* increment in Levenshtein similarity (95% *CrI* = [`r percent(coefs$lv_std_group_en_gvs_spa["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv_std_group_en_gvs_spa["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_lv_group1)`). English natives translating Catalan words were `r percent(abs(coefs$group_engmca_tvs_engmspa["estimate"]), accuracy = 0.01)` more accurate than those translating Spanish words (95% *CrI* = [`r percent(coefs$group_engmca_tvs_engmspa["q2_5"], accuracy = 0.01)`, `r percent(coefs$group_engmca_tvs_engmspa["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_group2)`). For every one *SD* increment in Levenshtein similarity, the effect increased in `r percent(abs(coefs$lv_std_group_engmca_tvs_engmspa["estimate"]), accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$lv_std_group_engmca_tvs_engmspa["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv_std_group_engmca_tvs_engmspa["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_lv_group1)`).

@@

m <- posterior_epreds_fixed %>% 
  mutate_at(vars(pthn_std, frequency_zipf_std), function(x) paste0(x, " SD")) %>% 
  mutate(
    pthn_std = case_when(
      pthn_std=="-1 SD" ~ "Low PTHN",
      pthn_std=="0 SD" ~ "Mean PTHN",
      pthn_std=="1 SD" ~ "High PTHN"
    ),
    frequency_zipf_std = case_when(
      frequency_zipf_std=="-1 SD" ~ "Low frequency",
      frequency_zipf_std=="0 SD" ~ "Mean frequency",
      frequency_zipf_std=="1 SD" ~ "High frequency"
    )
  )

ggplot(m, aes(lv_std, .epred, colour = pthn_std, fill = pthn_std)) +
  facet_wrap(~group) +
  geom_hline(yintercept = 0.5, colour = "grey") +
  geom_line(aes(group = interaction(pthn_std, .draw)), alpha = 0.25, size = 0.5, show.legend = TRUE) +
  stat_summary(aes(linetype = pthn_std), fun = mean, geom = "line", colour = "black", size = 0.75, show.legend = TRUE) +
  annotate("segment", x = -0.5, xend = -0.9, y = 0.85, yend = 0.85, arrow = arrow(length = unit(0.1, "cm")), size = .5) +
  annotate("segment", x = 2.25, xend = 2.65, y = 0.85, yend = 0.85, arrow = arrow(length = unit(0.1, "cm")), size = .5) +
  annotate("text", x = -0.5, y = 0.95, size = 2.75, label = "Less similar") +
  annotate("text", x = 2.15, y = 0.95, size = 2.75, label = "More similar") +
  labs(x = "Phonological similarity\n(Levenshtein, SD)", y = "P(Correct)", 
       colour = "PTHN", fill = "PTHN", linetype = "PTHN",
       title = "Population-level predictions") +
  scale_color_manual(values = c("#1E88E5", "#ff2976")) +
  scale_y_continuous(labels = function(x) percent(round(x, 2)), limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
  theme(
    legend.position = "top",
    plot.title = element_blank()
  )

@@

The conditional intra-class correlation (*ICC*) for items was `r percent(icc_conditional, accuracy = 0.01)`, indicating that responses from different participants to the same item were, on average, more similar to each other than responses to other items [@gelman2006data]. Item-level intercepts showed remarkable variability (*SD* = `r percent(sds$sd_word_intercept$.value, accuracy = 0.01)`, 95% CrI = [`r percent(sds$sd_word_intercept$.lower, accuracy = 0.01)`, `r percent(sds$sd_word_intercept$.upper, accuracy = 0.01)`]), indicating that participants' accuracy differed substantially across words: some words were correctly translated nearly always (e.g., Spanish *tomate*-*tomato*, /tomate/-/təmɑːtoʊ/, `r percent(word_example$tomate$estimate, accuracy = 0.01)`, 95% CrI = [`r percent(word_example$tomate$q2_5, accuracy = 0.01)`, `r percent(word_example$tomate$q97_5, accuracy = 0.01)`]), while others were almost never translated correctly (e.g., Catalan *colom*-*pigeon*, /kulom/-/pɪdʒɪn/, `r percent(word_example$colom$estimate, accuracy = 0.01)`, 95% CrI = [`r percent(word_example$colom$q2_5, accuracy = 0.01)`, `r percent(word_example$colom$q97_5, accuracy = 0.01)`]). Participant-level intercepts also showed high variability (*SD* = `r percent(sds$sd_participant_intercept$.value, accuracy = 0.01)`, 95% CrI = [`r percent(sds$sd_participant_intercept$.lower, accuracy = 0.01)`, `r percent(sds$sd_participant_intercept$.upper, accuracy = 0.01)`]), suggesting that average translation accuracy also varied significantly across participants. Finally, participant-level estimated slopes showed less variability: *SD* of $Frequency$, $PTHN$, $Levenshtein$, $PTHN \times Levenshtein$, and $Levenshtein \times Group 2$ ranged between `r percent(participant_ranef_range[1], accuracy = 0.01)`, and `r percent(participant_ranef_range[2], accuracy = 0.01)`. The participant-level estimated slopes of the $Leveshtein \times Group 1$ interaction are an exception, though (*SD* = `r percent(sds$sd_participant_lv_std_group_en_gvs_spa$.value, accuracy = 0.01)`, 95% CrI = [`r percent(sds$sd_participant_lv_std_group_en_gvs_spa$.lower, accuracy = 0.01)`, `r percent(sds$sd_participant_lv_std_group_en_gvs_spaE$.upper, accuracy = 0.01)`]). This indicates that the impact of phonological similarity ($Levenshtein$) on the difference in accuracy between English and Spanish natives differed substantially across participants.

Finally, participant-level effects showed weak Pearson correlations overall (<10%), with three exceptions. The estimated correlation between the intercepts and the slopes of $Frequency$ was `r percent(corr_data$intercept$estimate_frequency_zipf_std, accuracy = 0.01)` (95% CrI = [`r percent(corr_data$intercept$q2_5_frequency_zipf_std, accuracy = 0.01)`, `r percent(corr_data$intercept$q97_5_frequency_zipf_std, accuracy = 0.01)`], $BF_{Cor \neq 0}$ = `r round(h_cor_intercept_frequency, 2)`). The estimated correlation between the slopes of $Frequency$ and $Levenshtein \times Group 1$ was `r percent(corr_data$frequency_zipf_std$estimate_lv_std_group_en_gvs_spa, accuracy = 0.01)` (95% CrI = [`r percent(corr_data$frequency_zipf_std$q2_5_lv_std_group_en_gvs_spa, accuracy = 0.01)`, `r percent(corr_data$frequency_zipf_std$q97_5_lv_std_group_en_gvs_spa, accuracy = 0.01)`], $BF_{Cor \neq 0}$ = `r round(h_cor_frequency_zipf_lv_group1, 2)`). The estimated correlation between the intercepts and the slopes of $PTHN$ was `r percent(corr_data$intercept$estimate_pthn_std, accuracy = 0.01)` (95% CrI = [`r percent(corr_data$intercept$q2_5_pthn_std, accuracy = 0.01)`, `r percent(corr_data$intercept$q97_5_pthn_std, accuracy = 0.01)`], $BF_{Cor \neq 0}$ = `r round(h_cor_intercept_pthn, 2)`). The large uncertainty associated with these estimates, however, prevents us from interpreting the size of these correlations as practically meaningful.
